Precision calculations of cosmic shear projections

### 2017-02-10 conclusions.tex

For future surveys, the decrease in statistical errors places higher requirements on the accuracy of the theoretical modelling.    There is also, however, the need to be able to rapidly sample the multi-dimensional cosmological parameter likelihoods.  This requirement for computational speed is incompatible with a theoretical analysis that calculates a full spherical solution for the shear power spectrum, without adopting any approximation.  We therefore present alternative solutions, revisiting the work of  \citet{2012PhRvD..86b3001B} who showed that adopting the second-order extended Limber approximation of \citet{2008PhRvD..78l3506L} provides a representation of full spherical solution for the shear power spectrum that is accurate at the sub-percent level for $\ell > 3$.    We have verified this result and provide to the community our fast numerical implementation of all the approximations studied in this analysis,
and the slow calculation of the full projection within the publicly available package \textsc{nicaea} at \texttt{http://www.cosmostat.org/software/nicaea}.

Finally we propose that future surveys seek to optimise the statistical analyses of their cosmic shear data.  For example moving from the standard two-point shear correlation function statistic to the more stringent `COSEBI' statistic \citep{COSEBIs} renders the cosmic shear measurement insensitive to the low-$\ell$ scales where the Limber and flat-sky approximations make any impact on the precision of the theoretical modelling.

>>>>

For future surveys, the decrease in statistical errors places higher requirements on the accuracy of the theoretical modelling.    There is also, however, the need to be able to rapidly sample the multi-dimensional cosmological parameter likelihoods.  This requirement for computational speed is incompatible with a theoretical analysis that calculates a full spherical solution for the shear power spectrum, without adopting any approximation.  We therefore present alternative solutions, revisiting the work of  \citet{2012PhRvD..86b3001B} who showed that adopting the second-order extended Limber approximation of \citet{2008PhRvD..78l3506L} provides a representation of the full spherical solution for the shear power spectrum that is accurate at the sub-percent level for $\ell > 3$.    We have verified this result and provide to the community our fast numerical implementation of all the approximations studied in this analysis,
and the slow calculation of the full projection within the publicly available package \textsc{nicaea} at \texttt{http://www.cosmostat.org/software/nicaea}.

Finally we propose that future surveys seek to optimise the statistical analyses of their cosmic shear data.  For example moving from the standard two-point shear correlation function statistic to the more stringent `COSEBI' statistic \citep{COSEBIs} renders the cosmic shear measurement insensitive to the low-$\ell$ scales where the Limber and flat-sky approximations have an impact on the precision of the theoretical modelling.

### 2017-02-10 discussion.tex

In this paper we provide a solution in the form of the second-order extended
Limber approximation, but another option to consider is the used of alternative two-point
shear statistics that are less sensitive to accuracy in shear power spectrum
measurement at low $\ell$.   Both the aperture-mass dispersion, $\langle M_
{\rm ap}^2 \rangle$ \citep{1998MNRAS.296..873S}, and the Complete Orthogonal
Sets of E/B-mode Integrals (COSEBIs), $E_n$ \citep{COSEBIs} statistics satisfy
this requirement and are linearly related to the shear power spectrum in the
flat-sky approximation via integrals of the form 
%
\begin{align}
  \langle M_ {\rm ap}^2 \rangle(\theta) & = \int_0^{\infty}\frac{\d \ell \ell}{2\pi}
  \hat U^2(\theta\ell) P^\gamma(\ell),
  \label{eqn:integ}
  \\
  E_n & = \int_0^{\infty}\frac{\d \ell \ell}{2\pi} W_n(\ell) P^\gamma(\ell),
\end{align}
%
where the Fourier-space filter functions $\hat U$ and $W_n$ are defined in
\cite{1998MNRAS.296..873S} and \cite{COSEBIs}, respectively.
Figure~\ref{fig:filters} shows the integrands of these statistics, again
normalised to their maximum value, where the integrands are of the form $\ell
F(\ell) P^\gamma(\ell)$.  The lower middle panel in Figure~\ref{fig:filters}
shows the COSEBIs integrands for two angular ranges, $[1',100']$ and
$[0.8',350']$, where we only show the integrands for the lowest COSEBIs mode,
$E_1$, as the higher modes generally probe larger $\ell$-modes.  The lowest
panel shows the integrands of aperture mass dispersion statistics, for the same
two maximum angular ranges.

>>>>

In this paper we provide a solution in the form of the second-order extended
Limber approximation, but another option to consider is the use of alternative two-point
shear statistics that are less sensitive to accuracy in shear power spectrum
measurement at low $\ell$.   Both the aperture-mass dispersion, $\langle M_
{\rm ap}^2 \rangle$ \citep{1998MNRAS.296..873S}, and the Complete Orthogonal
Sets of E/B-mode Integrals (COSEBIs), $E_n$ \citep{COSEBIs} statistics satisfy
this requirement and are linearly related to the shear power spectrum in the
flat-sky approximation via integrals of the form 
%
\begin{align}
  \langle M_ {\rm ap}^2 \rangle(\theta) & = \int_0^{\infty}\frac{\d \ell \ell}{2\pi}
  \hat U^2(\theta\ell) P^\gamma(\ell),
  \label{eqn:integ}
  \\
  E_n & = \int_0^{\infty}\frac{\d \ell \ell}{2\pi} W_n(\ell) P^\gamma(\ell),
\end{align}
%
where the Fourier-space filter functions $\hat U$ and $W_n$ are defined in
\cite{1998MNRAS.296..873S} and \cite{COSEBIs}, respectively.
Figure~\ref{fig:filters} shows the integrands of these statistics, again
normalised to their maximum value, where the integrands are of the form $\ell
F(\ell) P^\gamma(\ell)$.  The lower middle panel in Figure~\ref{fig:filters}
shows the COSEBIs integrands for two angular ranges, $[1',100']$ and
$[0.8',350']$, where we only show the integrands for the lowest COSEBIs mode,
$E_1$, as the higher modes generally probe larger $\ell$-modes.  The lowest
panel shows the integrands of aperture mass dispersion statistics, for the same
two maximum angular ranges.
